---
title: ""
output: word_document
font-size: 11pt
---

# Load Libraries and Import Data into R  

```{r, message=FALSE, warning=FALSE}

library(tidyverse)
election_2016 <- read_csv("election_2016.csv")
head(election_2016)

```


\newpage
# Create Trump Vote Margins

```{r}

election_2016_margin <- election_2016 |> 
  mutate(
    trump_percent = (trump/totalvotes)*100,
    clinton_percent = (clinton/totalvotes)*100,
    trump_margin = trump_percent - clinton_percent
  )

  
head(election_2016_margin[,c(-5,-7,-6,-8)]) 

```


\newpage

# Summary Statistics

```{r}

summary(election_2016_margin)

```


Trump's Average Vote margin:    `r summary(election_2016_margin)[4,11]`

Median Poverty Rate:    `r summary(election_2016_margin)[3,7]`

People in least populous county:    `r summary(election_2016_margin)[1,8]`

\newpage

# Histogram of Trump's Percentage Vote Margin

```{r, message=FALSE}
trump_hist <- election_2016_margin |> 
  ggplot(aes(x = trump_margin)) +
  geom_histogram(fill = "blue", colour = "black", alpha = .4) + 
  scale_x_continuous(labels = scales::label_percent(scale = 1) ) +
  labs(title = "Histogram of Trump's Percentage Vote Margin",
       x = "Percent Margin",
       caption = "https://github.com/akhapwoyaco"
       ) +
  theme_light() + 
  theme(plot.title = element_text(hjust = 0.5) )

trump_hist

ggsave(plot = trump_hist, filename = "trump_hist.png", width = 35, height = 25, units = "cm", dpi = 400)
```


\newpage

# State Poverty Rates

```{r}

election_2016_poverty <- election_2016 |> 
  mutate( totalpov = (poverty/100) * population )

knitr::kable( head(election_2016_poverty[,c(-3,-4,-5,-6)]) )

```



```{r, message=FALSE}

election_2016_state <- election_2016_poverty |> group_by(stname) |> 
  summarize(totalpov_state = sum(totalpov),
            population_state = sum(population)) |> 
  mutate(statepovrate = totalpov_state/population_state ) |> 
  arrange(desc(population_state) )

knitr::kable( election_2016_state ) 
```

\newpage 

# Scatter PlotL State Poverty Rate (Percentage)

```{r}

povrate_scat <- election_2016_state |> 
  ggplot(aes(y = reorder(stname, statepovrate), x = statepovrate)) +
  geom_point() + 
  labs(
    x = "State Poverty Rate (Percentage)",
    caption = "https://github.com/akhapwoyaco") + 
  scale_x_continuous(
    labels = scales::label_percent(scale = 100) ) + 
  geom_text(
    data = subset(election_2016_state, population_state > 10000000),
    aes(label = population_state), hjust = -0.4, size = 3.5) +
  theme_bw() + 
  theme(
    axis.title.y = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_line(colour = "grey60" , linetype = "dashed")
  ) 

povrate_scat

ggsave(
  plot = povrate_scat, filename = "povrate_scatter.png",
  width = 40, height = 35, units = "cm", dpi = 400)
```

\newpage

We can evidently see a cluster of the top ten high populated states around the mid-range area of the poverty rate distribution of all states. We cannot entirely say that the most populous states have the higher rates, but their rates are high. States say, Mississippi, Louisiana, New Mexico and Kentucky have higher rates of all the states while their population is fairly low to average generally while the likes of New Hampshire, Hawaii have low populations as well as lower poverty rates comparatively. 


