---
title: "Logistic Regression 2"
author: "ACO"
subtitle: ""
date: "`r Sys.time()`"
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    toc: true
    toc_depth: 4
    toc_float: true
    code_folding: hide
---

<br>

# Analysis

## Data Import

```{r, warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines'}
library(readr)
library(ggplot2)
```


```{r, warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines'}
data_45 <- read_csv(
  "data1.csv")
#
data_45$Sex <- factor(data_45$Sex)
data_45$Improved <- ifelse(data_45$Improved==TRUE,1,0)
```

<br>

## Logistic Regressions Tables

### `Improved ~ Sex` logistic model.

```{r,warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines'} 
#Sex Only
imp_sex_log <- glm(Improved ~ Sex, data = data_45, family = "binomial")
summary(imp_sex_log)
```

<br>

### `Improved ~ Height` logistic model.

```{r,warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines'} 
#Height Only
imp_height_log <- glm(
  Improved ~ `Height (cm)`, 
  data = data_45, family = "binomial")
summary(imp_height_log)
```

<br>

### `Improved ~ Weight` logistic model.

```{r,warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines'} 
#Weight Only
imp_wei_log <- glm(
  Improved ~ `Weight (kg)`, 
  data = data_45, family = "binomial")
summary(imp_wei_log)

```

<br>

### `Improved ~ Sex + Height + Weight` logistic model.

```{r,warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines'} 
#combination
imp_hei_wei_log <- glm(
  Improved ~ Sex + `Height (cm)` + `Weight (kg)`, 
  data = data_45, family = "binomial")
summary(imp_hei_wei_log)
```

<br>

### `Improved ~ Sex + Height + Weight + Height*Weight` logistic model.

```{r,warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines'} 
#Combined + weight height interaction
imp_combi_log <- glm(
  Improved ~ Sex + `Height (cm)`*`Weight (kg)`, 
  data = data_45, family = "binomial")
summary(imp_combi_log)
```

<br>

1. Sex is a significant predictor of Improved 
($\beta$= `r summary(imp_sex_log)$coefficients[2,1]`,$p$=`r summary(imp_sex_log)$coefficients[2,4]`).

2. Males show an increase in the Odds of the subject's symptoms by `r 100 * (exp(summary(imp_combi_log)$coefficients[2,1]) - 1)` %.
3. When looking at Weight , Height , Sex , and the interaction of Weight and Height the below terms are significant. Note: non-significant terms were removed.

+ Weight ($\beta$= `r summary(imp_combi_log)$coefficients[4,1]`, $p$= `r summary(imp_combi_log)$coefficients[4,4]`)
+ Height ($\beta$= `r summary(imp_combi_log)$coefficients[3,1]`, $p$= `r summary(imp_combi_log)$coefficients[3,4]`)
+ Sex ($\beta$= `r summary(imp_combi_log)$coefficients[2,1]`, $p$= `r summary(imp_combi_log)$coefficients[2,4]`)
+ Weight:Height ($\beta$= `r summary(imp_combi_log)$coefficients[5,1]`, $p$= `r summary(imp_combi_log)$coefficients[5,4]`)

4. When comparing the two models from last week and the two from this week on AIC, 
the [Weight only]/[ Height only]/[ Sex only]/[combination] model is the best.

<br>

# Figures

## Improved vs Height

```{r,warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines', fig.align='center', dpi=350, fig.width=10, fig.height=8} 
hei_impr_plot <- data_45 |>
  ggplot(aes(x = `Height (cm)`, y = Improved)) +
  geom_jitter(width = 0, height = 0.02) +
  geom_smooth(
    method="glm", se=T, 
              method.args = list(family="binomial"))+ 
  labs(x = "Height", caption = "https://github.com/akhapwoyaco") +
  theme_bw()

hei_impr_plot
ggsave("hei_impr_plot.png",plot = hei_impr_plot, 
       width = 35, height = 20, unit = "cm", dpi = 400)
```

<br>

## Improved vs Weight

```{r,warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines', fig.align='center', dpi=350, fig.width=10, fig.height=8} 
wei_impr_plot <- data_45  |> 
  ggplot(aes(x = `Weight (kg)`, y = Improved)) +
  geom_jitter(width = 0, height = 0.02) +
  geom_smooth(method="glm", se=T, 
              method.args = list(family=binomial))+ 
  labs(x = "Weight", caption = "https://github.com/akhapwoyaco") +
  theme_bw()
wei_impr_plot
ggsave("wei_impr_plot.png",plot = wei_impr_plot, 
       width = 35, height = 20, unit = "cm", dpi = 400)
```
<br>

## Improved vs Height by Sex

```{r,warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines', fig.align='center', dpi=350, fig.width=10, fig.height=8} 
hei_impr_sex_plot <- data_45 |> 
  ggplot(aes(x = `Height (cm)`, y = Improved, color = Sex)) +
  geom_jitter(width = 0, height = 0.02) +
  geom_smooth(method="glm", se=T, 
              method.args = list(family=binomial))+ 
  scale_color_discrete(labels = c("Female","Male")) + 
  labs(x = "Height", caption = "https://github.com/akhapwoyaco") +
  theme_bw()
hei_impr_sex_plot
ggsave("hei_impr_sex_plot.png",plot = hei_impr_sex_plot, 
       width = 35, height = 20, unit = "cm", dpi = 400)
```

<br>

## Improved vs Weight by Sex

```{r,warning=FALSE,message=FALSE,attr.source='.numberLines',attr.output='.numberLines', fig.align='center', dpi=350, fig.width=10, fig.height=8} 
wei_impr_sex_plot <- data_45  |> 
  ggplot(aes(x = `Weight (kg)`, y = Improved, color = Sex)) +
  geom_jitter(aes(color=Sex),width = 0, height = 0.02) +
  geom_smooth(method="glm", se=T, 
              method.args = list(family=binomial)) + 
  scale_color_discrete(labels = c("Female","Male")) + 
  labs(x = "Weight", caption = "https://github.com/akhapwoyaco") +
  theme_bw()
wei_impr_sex_plot
ggsave("wei_impr_sex_plot.png",plot = wei_impr_sex_plot, 
       width = 35, height = 20, unit = "cm", dpi = 400)
```

<br>